TODO: A Survey of StarCraft AI Techniques

### 2012-05-14 survey.tex
Pass over related work

{\color{blue}
\begin{LARGE}
TODO merge above and bottom\\
\end{LARGE}
\textbf{Planning}
TODO (Santi?) just a paragraph on exposing why there is a planning problem in RTS AI.

@@

\textbf{Learning}\\
When talking about learning in RTS games, with regard to a particular match against an opponent, one has to differentiate between:
\begin{itemize}
\item prior learning: done before the match, like when a bot learns what is likely to happen on a given match-up (and map) from replays, or when a bot optimizes its parameters beforehand through simulated annealing or genetic evolution.
\item in-game learning: done between the different actions (decisions) taken during a game. For instance, a bot can take into account the successes and failures of its attacks (location, type, timings).
\item in-match learning: done between the different games of match, it encompasses opponent modeling across multiple games and adapting the bots' strategy to its successes and failures.
\end{itemize}
All these learning can be supervised (labeled replays, scripted situations) or unsupervised (unlabeled data, exploratory situations). The last two kinds of situations of learning are more inclined towards reinforcement learning because the AI has to deal with the exploration-exploitation duality. %prior learning: supervised \cite{WeberCIG09} \cite{SynnaeveCIG11}, unsupervised: \cite{SynnaeveAIIDE11}

@@

\subsection{RTS AI Decomposition}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and their properties: uncertainty (from intentions and from partial observation) is going higher as the abstraction levels are raising. The timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft. Spatial and temporal reasoning are indicated for part for which greedy solutions are really not efficient.}
    \label{fig:levels-abstraction}
\end{figure*}
As a work on RTS AI has to address several of these aforementioned problems together, we decided to study decompose RTS AI in three tasks, widely known by RTS players: strategy, tactics and units control (micro-management), see Figure~\ref{fig:levels-abstraction}. We consider strategy to be everything related to the \textit{technology tree}, being the buildings and the upgrades, and the army composition. It is the most intentional level, as a player performs a strategy with future stances (aggressive, defensive, economy, technology) and tactics in mind. To change its strategy, we consider that a player need about 3 minutes: they should need to construct a few new buildings and/or research some upgrades and/or produce units changing the army composition. Tactics represent the attacks and defenses. For the human player, decisions on tactics are made from spatial (exploiting the terrain) and temporal (army movements) reasoning, constrained on the possible types of attacks by the army composition of the player and their opponent. Units control, also called micro-management, is how the player takes action to maximize the efficiency of its units (applying the tactics) in real-time (simultaneous) and adversary conditions. As shown in Figure~\ref{fig:levels-abstraction}, strategy constrains possible tactics which in turn constrain units control. On a prediction/adaptation point of view, the units observation level gives information about the opponent's tactics and strategy. Flaws in opponents tactics can also steer our strategy into exploiting them.\\
{\color{blue}
perhaps TODO examples?
}

@@

In the industry, strategies are often ``hard coded'' with finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003}. The idea behind FSM is to decompose an object's behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Hierarchical FSM give higher (strategic) states some control over lower oned (actions). However, strategic decisions have to be made in a context, not in a vacuum. It is hard to encode dynamic, adaptive behaviors in (H)FSM due to the exponential number of transitions (between states) to account for. To this end, several approaches have been developed like hierarchical task networks, behavior trees and the comeback of STRIPS \cite{FikesSTRIPS} planning in the industry \cite{orkinGDC_FEAR}. Regardless, one cannot but notice that strategy adaptation is not the strong point of industrial RTS AI. Strategy adaptation is part of what is called opponent modeling. An adapting robot needs to keep track of what the enemy is doing in order to estimate the probability that they will use a specific strategy and adapt its own strategy to it.

In research, the main methods used for strategic decision-making are case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}. Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). Onta\~{n}\'{o}n et al. \cite{CBR_Planning} base their real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR learned from expert demonstrations on Wargus. They update their previous CBP approach by using situation assessment for retrieving the ``right'' plan, and improve the speed of CBP by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). Hoang et al. \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. %Madeira et al. \cite{Madeira06}

@@

Most commercial games avoid gathering this information by using perfect information all the time (cheating). Sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. {\color{blue}XXX (perhaps move it above, to the start of subsection}

@@

\subsection{Tactics}
Units have different abilities, which leads to different possible tactics. Most often, fast or invisible or flying units are not cost-effective in head-to-head fights against slower bulky units. Taking good tactical decisions is taking advantage of the terrain and positions of the different units (in a timely manner). Terrain analysis supplies the AI with structured information about the map in order to help making decisions. This analysis is usually performed off-line, in order to save CPU time during the game. Pottinger \cite{Pottinger00} described the \emph{BANG} engine implemented by Ensemble Studios for the game Age of Empires II. This engine provides terrain analysis functionalities to the game using influence maps and areas with connectivity information. Forbus et al. \cite{Forbus2002} showed the importance to have qualitative spatial information for wargames, for which they used geometric and pathfinding analysis. Hale et al. \cite{Hale08} presented a 2D geometric navigation mesh generation method from expanding convex regions from seeds. Finally, Perkins \cite{Perkins10} applied Voronoi decomposition (then pruning) %\cite{Karavelas04}
to detect regions and relevant choke points in RTS maps. This approach is implemented for StarCraft in BWTA\footnote{\url{http://code.google.com/p/bwta/}}.
% Isla 2006 Game AI Programming Wisdom 3. Charles River Media. chapter Probabilistic Target Tracking and Search Using Occupancy Maps, 379–388. ?

Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in Tankbattle. Weber et al. \cite{WeberAIIDE11} used a particle model with a simple linear trajectory update to track opponent units under fog of war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{Chung05} adapted Monte-Carlo tree search (MCTS) to planning in RTS games and applied it to a capture-the-flag mod of Open RTS. Balla and Fern \cite{UCT} applied upper confidence bounds on trees (UCT: a MCTS algorithm) to tactical assault planning in Wargus. On influence maps, \cite{teamCompositionRTS} studied team composition and maneuvering by learning a self-organizing map, while \cite{HagelbackJ08} presented a multi-agent potential field based bot. Finally, Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial (tactical) reasoning in RTS games.

\subsection{Units Control}
Units control, also called micro-management in RTS games, aims at maximizing the effectiveness of units, it starts with pathfinding and goes up to simultaneous control of units of different types in complex battles on heterogeneous terrain. The most common pathfinding algorithm is A*, but its big problem is CPU time and memory consumption, hard to satisfy in a complex, dynamic, real-time environment on multiple units. %To solve this the search space needs to be reduced as much as possible. 
A common technique consists in reducing the search into a navigation graph map. Among others, Triangulation Reduction A* computes polygonal triangulations on a grid-based map \cite{Demyen_2006}. One of the problems of A* related approaches in RTS games is that moving units block/get in the way of/ each others and the path has to be recomputed often. Craig Reynolds \cite{Reynolds_1999} introduced algorithmic steering behaviors, which simulate the aggregate motion of swarms. His model describes how a system of multiple agents, each steering according to simple local rules, can produce the collective movement found in flocks of birds. Danielsiek et al. \cite{Danielsiek_2008} studied intelligent squad movements to do tactical moves to flank the opponent in a RTS game this way. An alternative to flocking is to use reciprocal velocity obstacles (RVO) \cite{Berg_2008}, since it is easier to control the behavior and they guarantee to prevent collisions. With this algorithm, we can move a large squad crossing another squad and avoid all possible collisions between units. {\color{blue}Do we put RVO? As I don't remember it being used in any game or game AI research.} In recent commercial RTS games like Starcraft 2 or Supreme Commander 2, flocking-like behaviors are inspired of continuum crowds (``flow field'') \cite{Treuille2006}, and a comprehensive review about (grid-based) pathfinding was recently done by Sturtevant \cite{sturtevant2012benchmarks}.

>>>>

{\color{red}
TODO merge above and bottom
}

{\color{blue}
\textbf{Planning}
TODO (Santi?) just a paragraph on exposing why there is a planning problem in RTS AI.

@@

\textbf{Learning}\\
When talking about learning in RTS games, with regard to a particular match against an opponent (where a match is a sequence of games against the same opponent), one has to differentiate between:
\begin{itemize}
\item Prior learning: done before the match, like when a bot learns what is likely to happen on a given match-up (and map) from replays, or when a bot optimizes its parameters beforehand through simulated annealing or genetic evolution.
\item In-game learning: done between the different actions (decisions) taken during a game. For instance, a bot can take into account the successes and failures of its attacks (location, type, timings).
\item In-match learning: done between the different games of the match, it encompasses opponent modeling across multiple games and adapting the bots' strategy to its successes and failures.
\end{itemize}
Learning in all these situations can be supervised (labeled replays, scripted situations) or unsupervised (unlabeled data, exploratory situations). The last two kinds of situations of learning are more inclined towards reinforcement learning because the AI has to deal with the exploration-exploitation duality. %prior learning: supervised \cite{WeberCIG09} \cite{SynnaeveCIG11}, unsupervised: \cite{SynnaeveAIIDE11}

@@

\subsection{RTS AI Decomposition}
\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/levels_abstraction.pdf}
    \caption{RTS AI levels of abstraction and their properties: uncertainty (from intentions and from partial observation) is going higher as the abstraction levels are raising. The timings on the right correspond to an estimate of the duration of a behavior switch in StarCraft. Spatial and temporal reasoning are indicated for part for which greedy solutions are really not efficient.}
    \label{fig:levels-abstraction}
\end{figure}
As work on RTS AI has to address several of the aforementioned problems together, in this section, we divide the existing work on RTS AI in three different levels of abstraction, widely known by RTS players: strategy, tactics and units control (micro-management), see Figure~\ref{fig:levels-abstraction}. We consider strategy to be everything related to the \textit{technology tree}, buildings, upgrades, and army composition. It is the most deliberative level, as a player performs a strategy with future stances (aggressive, defensive, economy, technology) and tactics in mind. To change its strategy, we consider that a player need about 3 minutes: they should need to construct a few new buildings and/or research some upgrades and/or produce units changing the army composition. Tactics represent the attacks and defenses between groups of units. For the human player, decisions on tactics are made from spatial (exploiting the terrain) and temporal (army movements) reasoning, constrained on the possible types of attacks by the army composition of the player and their opponent. Unit control, also called micro-management, is how the player takes action to maximize the efficiency of its individual units (applying the tactics) in real-time (simultaneous) and adversary conditions. As shown in Figure~\ref{fig:levels-abstraction}, strategy constrains possible tactics which in turn constrain units control. On a prediction/adaptation point of view, the units observation level gives information about the opponent's tactics and strategy. Flaws in opponents tactics can also steer our strategy into exploiting them.\\
{\color{blue}
perhaps TODO examples?
}

@@

In commercial RTS games, strategies are often ``hard coded'' with finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003}. The idea behind FSMs is to decompose an object's behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Hierarchical FSM give higher (strategic) states some control over lower ones (actions). However, strategic decisions have to be made in a context, not in a vacuum. It is hard to encode dynamic, adaptive behaviors in (H)FSM due to the exponential number of transitions (between states) to account for. To this end, several approaches have been developed like hierarchical task networks, behavior trees and the comeback of STRIPS \cite{FikesSTRIPS} planning in the industry \cite{orkinGDC_FEAR}. Regardless, one cannot but notice that adaptability is not the strong point of industrial RTS AI. In the context of RTS AI strategy adaptation is related to the idea of opponent modeling. An adapting AI needs to keep track of what the enemy is doing in order to estimate the probability that they will use a specific strategy and adapt its own strategy accordingly.

In academic RTS AI research, the main methods used for strategic decision-making have been case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}. Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). Onta\~{n}\'{o}n et al. \cite{CBR_Planning} base real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR learned from expert demonstrations on Wargus. They update their previous CBP approach by using situation assessment for retrieving the ``right'' plan, and improve the speed of CBP by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). Hoang et al. \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. %Madeira et al. \cite{Madeira06}

@@

One final consideration is that RTS games are typically partially observable. For example games like StarCraft implement the ``fog of war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Therefore, in order to play an RTS game, players need to scout the opponent in order to obtain information about the opponent's strategy. Most commercial games avoid this problem by using perfect information all the time (cheating). In order to make the human player believe the built-in AI of this games does scouting, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. %{\color{blue}XXX (perhaps move it above, to the start of subsection}

@@

\subsection{Tactics}
Units have different abilities, which leads to different possible tactics. For example, very often, fast, invisible or flying units are not cost-effective in head-to-head fights against slower bulky units. Taking good tactical decisions is taking advantage of the terrain and positions of the different units (in a timely manner). Terrain analysis supplies the AI with structured information about the map in order to help making decisions. This analysis is usually performed off-line, in order to save CPU time during the game. Pottinger \cite{Pottinger00} described the \emph{BANG} engine implemented by Ensemble Studios for the game Age of Empires II. This engine provides terrain analysis functionalities to the game using influence maps and areas with connectivity information. Forbus et al. \cite{Forbus2002} showed the importance to have qualitative spatial information for wargames, for which they used geometric and pathfinding analysis. Hale et al. \cite{Hale08} presented a 2D geometric navigation mesh generation method from expanding convex regions from seeds. Finally, Perkins \cite{Perkins10} applied Voronoi decomposition (then pruning) %\cite{Karavelas04}
to detect regions and relevant choke points in RTS maps. This approach is implemented for StarCraft in BWTA\footnote{\url{http://code.google.com/p/bwta/}}.
% Isla 2006 Game AI Programming Wisdom 3. Charles River Media. chapter Probabilistic Target Tracking and Search Using Occupancy Maps, 379–388. ?

Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog of war in Tankbattle. Weber et al. \cite{WeberAIIDE11} used a particle model with a simple linear trajectory update to track opponent units under fog of war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{Chung05} adapted Monte-Carlo tree search (MCTS) to planning in RTS games and applied it to a capture-the-flag mod of Open RTS. Balla and Fern \cite{UCT} applied upper confidence bounds on trees (UCT: a MCTS algorithm) to tactical assault planning in Wargus. % (Santi: I commented this one, since it's mentioned above) On influence maps, \cite{teamCompositionRTS} studied team composition and maneuvering by learning a self-organizing map, while \cite{HagelbackJ08} presented a multi-agent potential field based bot.  
Finally, Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial (tactical) reasoning in RTS games.

\subsection{Units Control}
Units control, also called micro-management in RTS games, aims at maximizing the effectiveness of units, it starts with pathfinding and goes up to simultaneous control of units of different types in complex battles on heterogeneous terrain. The most common pathfinding algorithm is A*, but its big problem is CPU time and memory consumption, hard to satisfy in a complex, dynamic, real-time environment on multiple units. %To solve this the search space needs to be reduced as much as possible. 
A common technique consists in reducing the search into a navigation graph map. Among others, Triangulation Reduction A* computes polygonal triangulations on a grid-based map \cite{Demyen_2006}. One of the problems of A* related approaches in RTS games is that moving units block/get in the way of/ each others and the path has to be recomputed often. Craig Reynolds \cite{Reynolds_1999} introduced algorithmic steering behaviors, which simulate the aggregate motion of swarms. His model describes how a system of multiple agents, each steering according to simple local rules, can produce the collective movement found in flocks of birds. Danielsiek et al. \cite{Danielsiek_2008} studied intelligent squad movements to do tactical moves to flank the opponent in a RTS game this way. %An alternative to flocking is to use reciprocal velocity obstacles (RVO) \cite{Berg_2008}, since it is easier to control the behavior and they guarantee to prevent collisions. With this algorithm, we can move a large squad crossing another squad and avoid all possible collisions between units. {\color{blue}Do we put RVO? As I don't remember it being used in any game or game AI research.} (Santi: I agree, let's not mention it)
In recent commercial RTS games like Starcraft 2 or Supreme Commander 2, flocking-like behaviors are inspired of continuum crowds (``flow field'') \cite{Treuille2006}. A comprehensive review about (grid-based) pathfinding was recently done by Sturtevant \cite{sturtevant2012benchmarks}.
